---
title: "Project Report"
format: html
editor: visual
---

Deasy Pasaribu

**Title: Bank Customer Churn**

Dataset Overview

The dataset contains information about 10,000 bank customers from 3 geography areas and long span of ages. The primary goal is to predict the 'Exited' column, indicating whether a customer has churned (1) or retained (0).

1.  Data Structure and Summary

    ```{r}
    # Load library
    library(tidyverse)

    customer_churn <- read.csv("Churn_Modelling.csv")
    str(customer_churn)
    summary(customer_churn)
    any(is.na(customer_churn))

    # Checking for missing value
    knitr::kable(sapply(customer_churn, function(x) sum(is.na(x))), col.names = c("Missing Value"))
    ```

2.  Data pre-processing

    ```{r}
    # Load library
    library(caret)

    # Remove insignificant variables
    customer_churn = customer_churn[, !names(customer_churn) %in% c('RowNumber', 'CustomerId', 'Surname')]

    # Convert as factor
    customer_churn$Geography = factor(customer_churn$Geography, labels=c(0, 1, 2))
    customer_churn$Gender = factor(customer_churn$Gender, labels=c(0, 1))

    # Ensure model compatibility
    customer_churn$Age = log(customer_churn$Age)
    customer_churn$CreditScore = log(customer_churn$CreditScore)
    customer_churn$Balance = log(customer_churn$Balance)
    customer_churn[customer_churn$Balance == -Inf, 'Balance'] <- 0

    # Set scale "0 to 1"
    fun_scale_0to1 <- function(x) {
      (x - min(x)) / (max(x) - min(x))
      }
    #customer_churn$Age = fun_scale_0to1(customer_churn$Age) #########
    customer_churn$CreditScore = fun_scale_0to1(customer_churn$CreditScore)
    customer_churn$Balance = fun_scale_0to1(customer_churn$Balance)
    customer_churn$EstimatedSalary = fun_scale_0to1(customer_churn$EstimatedSalary)
    head(customer_churn, 25)
    ```

3.  Logistic Regression

    ```{r}
    # Load libraries
    library(caret)
    library(MLmetrics)
    library(dplyr)

    # Load data
    customer_churn <- read.csv("Churn_Modelling.csv")

    # Remove identifier columns
    customer_churn <- customer_churn %>% select(-RowNumber, -CustomerId, -Surname)

    # Convert target to factor
    customer_churn$Exited <- as.factor(customer_churn$Exited)

    # Convert categorical variables to factor
    customer_churn$Gender <- as.factor(customer_churn$Gender)
    customer_churn$Geography <- as.factor(customer_churn$Geography)
    customer_churn$HasCrCard <- as.factor(customer_churn$HasCrCard)
    customer_churn$IsActiveMember <- as.factor(customer_churn$IsActiveMember)

    # Fit the regression model (full)
    churn_mod = glm(Exited ~ ., customer_churn, family = "binomial")
    summary(churn_mod)

    # Show odds ratios
    exp(coef(churn_mod))
    ```

    ```{r}
    # Model of Selected Variables
    # Model 1
    churn_mod1 <- glm(Exited ~ Age + IsActiveMember, customer_churn, family = "binomial")
    summary(churn_mod1)

    # Model 2
    churn_mod2 <- glm(Exited ~ Age + IsActiveMember + Balance, customer_churn, family = "binomial")
    summary(churn_mod2)

    # Model 3
    churn_mod3 <- glm(Exited ~ Age + IsActiveMember + Balance + Geography + Gender, customer_churn, family = "binomial")
    summary(churn_mod3)

    # Model 4
    churn_mod4 <- glm(Exited ~ Age + IsActiveMember + Balance + Geography + Gender + NumOfProducts, customer_churn, family = "binomial")
    summary(churn_mod4)

    # Model 5
    churn_mod5 <- glm(Exited ~ Age + IsActiveMember + Balance + Geography + Gender + NumOfProducts + CreditScore, customer_churn, family = "binomial")
    summary(churn_mod5)
    ```

    ```{r}
    # Evaluate baseline model (logistic regression) performance using F1-Score
    evaluate_model <- function(model, customer_churn) {
      prob <- predict(model, type = "response")
      pred <- ifelse(prob > 0.5, 1, 0)
      actual <- as.numeric(as.character(customer_churn$Exited))
      acc <- Accuracy(pred, actual)
      f1 <- F1_Score(pred, actual, positive = "1")
      return(list(accuracy = acc, f1_score = f1))}

    eval_churn_mod <- evaluate_model(churn_mod, customer_churn)
    print(eval_churn_mod)
    eval_mod1 <- evaluate_model(churn_mod1, customer_churn)
    print(eval_mod1)
    eval_mod2 <- evaluate_model(churn_mod2, customer_churn)
    print(eval_mod2)
    eval_mod3 <- evaluate_model(churn_mod3, customer_churn)
    print(eval_mod3)
    eval_mod4 <- evaluate_model(churn_mod4, customer_churn)
    print(eval_mod4)
    eval_mod5 <- evaluate_model(churn_mod5, customer_churn)
    print(eval_mod5)
    ```

4.  Multicollinearity

    ```{r}
    # Checking multicollinearity
    library(car)

    vif(churn_mod)
    vif(churn_mod1)
    vif(churn_mod2)
    vif(churn_mod3)
    vif(churn_mod4)
    vif(churn_mod5)
    ```

5.  Confidence Interval, Odds Ratio, and Wald Test

    ```{r}
    ## CI and Odds Ratio + CI ##
    # Full model
    confint(churn_mod)
    exp(cbind(OR = coef(churn_mod), confint(churn_mod)))

    # Model 1
    confint(churn_mod1)
    exp(cbind(OR = coef(churn_mod1), confint(churn_mod1)))

    # Model 2
    confint(churn_mod2)
    exp(cbind(OR = coef(churn_mod2), confint(churn_mod2)))

    # Model 3
    confint(churn_mod3)
    exp(cbind(OR = coef(churn_mod3), confint(churn_mod3)))

    # Model 4
    confint(churn_mod4)
    exp(cbind(OR = coef(churn_mod4), confint(churn_mod4)))

    # Model 5
    confint(churn_mod5)
    exp(cbind(OR = coef(churn_mod5), confint(churn_mod5)))
    ```

6.  Goodness of Fit

    ```{r}
    # Perform AIC
    AIC(churn_mod)
    AIC(churn_mod1)
    AIC(churn_mod2)
    AIC(churn_mod3)
    AIC(churn_mod4)
    AIC(churn_mod5)
    ```

7.  Likelihood Ratio Test

    ```{r}
    # Likelihood Ratio Test

    cat("#Likelihood Ratio for Model 1# \n")
    anova(churn_mod1, test="Chisq")
    cat("#Likelihood Ratio for Model 2# \n")
    anova(churn_mod2, test="Chisq")
    cat("#Likelihood Ratio for Model 3# \n")
    anova(churn_mod3, test="Chisq")
    cat("#Likelihood Ratio for Model 4# \n")
    anova(churn_mod4, test="Chisq")
    cat("#Likelihood Ratio for Model 5# \n")
    anova(churn_mod5, test="Chisq")
    cat("#Likelihood Ratio for Full model# \n")
    anova(churn_mod, test="Chisq")
    ```

8.  Hosmer-Lemeshow Test

    ```{r}
    # Load library
    library(ResourceSelection)

    # Convert factor to numeric
    customer_churn$Exited <- as.numeric(as.character(customer_churn$Exited))

    # Hosmer-Lemeshow Test
    cat("#Hosmer-Lemeshow for Model 1# \n")
    hoslem.test(customer_churn$Exited, fitted(churn_mod1))

    cat("#Hosmer-Lemeshow for Model 2# \n")
    hoslem.test(customer_churn$Exited, fitted(churn_mod2))

    cat("#Hosmer-Lemeshow for Model 3# \n")
    hoslem.test(customer_churn$Exited, fitted(churn_mod3))

    cat("#Hosmer-Lemeshow for Model 4# \n")
    hoslem.test(customer_churn$Exited, fitted(churn_mod4))

    cat("#Hosmer-Lemeshow for Model 5# \n")
    hoslem.test(customer_churn$Exited, fitted(churn_mod5))

    cat("#Hosmer-Lemeshow for Full Model \n")
    hoslem.test(customer_churn$Exited, fitted(churn_mod))
    ```

9.  Model Evaluation using Confusion Matrix

    ```{r}
    #Load library
    library(caret)

    ## Confusion Matrix ##
    # Full model
    prob_mod <- predict(churn_mod, type = "response")
    pred_mod <- ifelse(prob_mod > 0.5, 1, 0)
    conf_matrix_mod <- confusionMatrix(factor(pred_mod), factor(customer_churn$Exited), positive = "1")
    cat("Full Model:\n")
    print(conf_matrix_mod)

    # Model 1
    prob_mod1 <- predict(churn_mod1, type = "response")
    pred_mod1 <- ifelse(prob_mod1 > 0.5, 1, 0)
    conf_matrix_mod1 <- confusionMatrix(factor(pred_mod1), factor(customer_churn$Exited), positive = "1")
    cat("Model 1:\n")
    print(conf_matrix_mod1)

    # Model 2
    prob_mod2 <- predict(churn_mod2, type = "response")
    pred_mod2 <- ifelse(prob_mod2 > 0.5, 1, 0)
    conf_matrix_mod2 <- confusionMatrix(factor(pred_mod2), factor(customer_churn$Exited), positive = "1")
    cat("Model 2:\n")
    print(conf_matrix_mod2)

    # Model 3
    prob_mod3 <- predict(churn_mod3, type = "response")
    pred_mod3 <- ifelse(prob_mod3 > 0.5, 1, 0)
    conf_matrix_mod3 <- confusionMatrix(factor(pred_mod3), factor(customer_churn$Exited), positive = "1")
    cat("Model 3:\n")
    print(conf_matrix_mod3)

    # Model 4
    prob_mod4 <- predict(churn_mod4, type = "response")
    pred_mod4 <- ifelse(prob_mod4 > 0.5, 1, 0)
    conf_matrix_mod4 <- confusionMatrix(factor(pred_mod4), factor(customer_churn$Exited), positive = "1")
    cat("Model 4:\n")
    print(conf_matrix_mod4)

    # Model 5
    prob_mod5 <- predict(churn_mod5, type = "response")
    pred_mod5 <- ifelse(prob_mod5 > 0.5, 1, 0)
    conf_matrix_mod5 <- confusionMatrix(factor(pred_mod5), factor(customer_churn$Exited), positive = "1")
    cat("Model 5:\n")
    print(conf_matrix_mod5)
    ```

10. ROC-AUC

    ```{r}
    library(pROC)

    # Probability predicition
    churn_all_models <- list(
      Model1 = prob_mod1,
      Model2 = prob_mod2,
      Model3 = prob_mod3,
      Model4 = prob_mod4,
      Model5 = prob_mod5,
      FullModel = prob_mod)

    colors <- c("red", "green", "blue", "yellow", "orange", "black")

    # ROC
    rocs <- Map(function(pred) roc(customer_churn$Exited, pred), churn_all_models)

    # Plot ROC
    plot(rocs[[1]], col = colors[1], lwd = 2, main = "ROC Curves – All Models")
    invisible(mapply(function(r, col) lines(r, col = col, lwd = 2), rocs[-1], colors[-1]))
    legend("bottomright", legend = names(churn_all_models), col = colors, lwd = 2)

    ```

11. Cross Validation

    ```{r}
    # Load library
    library(caret)

    # Set up cross-validation
    cv_ctrl <- trainControl(method = "cv", number = 10, classProbs = TRUE, summaryFunction = twoClassSummary)

    # Convert target to factor for caret
    customer_churn$Exited <- factor(customer_churn$Exited, levels = c(0, 1), labels = c("No", "Yes"))

    # Train logistic regression with CV
    set.seed(123)
    logit_cv <- train(Exited ~ ., data = customer_churn,
                      method = "glm",
                      family = "binomial",
                      trControl = cv_ctrl,
                      metric = "ROC")  # optimize for AUC

    logit_cv
    ```

12. EDA: Target Variable Distribution

    ```{r}
    library(ggplot2)
    customer_churn %>%
      count(Exited) %>%
      mutate(percentage = n / sum(n) * 100)

    ggplot(customer_churn, aes(x = factor(Exited))) +
      geom_bar(fill = c("green", "red")) +
      labs(title = "Customer Churn Distribution", x = "Exited", y = "Count")
    ```

13. EDA: Demographic Summary

    ```{r}
    customer_churn %>% count(Gender)
    customer_churn %>%
      group_by(Gender) %>%
      summarise(min_age = min(Age),
                max_age = max(Age),
                mean_age = mean(Age),
                median_age = median(Age))
    customer_churn %>% count(Geography)
    ```

14. EDA: Categorical Variables & Churn Trends

    ```{r}
    library(tidyverse)
    library(ggplot2)
    # Churn rate by Geography
    churn_by_geography <- customer_churn %>%
      group_by(Geography) %>%
      summarise(churn_rate = mean(Exited) * 100)
    churn_by_geography

    # Churn rate by Gender
    churn_by_gender <- customer_churn %>%
      group_by(Gender) %>%
      summarise(churn_rate = mean(Exited) * 100)
    churn_by_gender

    # Plot (by Geography)
    ggplot(churn_by_geography, aes(x = Geography, y = churn_rate, fill = Geography)) +
      geom_col() +
      labs(title = "Churn Rate by Geography", y = "Churn Rate (%)") +
      theme_minimal()

    # Plot (by Gender)
    ggplot(churn_by_gender, aes(x = Gender, y = churn_rate, fill = Gender)) +
      geom_col() +
      labs(title = "Churn Rate by Gender", y = "Churn Rate (%)") +
      theme_minimal()
    ```

15. EDA: Numerical Variables & Churn Trends

    ```{r}
    # Churn rate summary
    numerical_summary <- customer_churn %>%
      group_by(Exited) %>%
      summarise(
        Age_mean = mean(Age),
        Age_median = median(Age),
        Age_min = min(Age),
        Age_max = max(Age),
        Balance_mean = mean(Balance),
        Balance_median = median(Balance),
        CreditScore_mean = mean(CreditScore),
        Tenure_mean = mean(Tenure),
        EstimatedSalary_mean = mean(EstimatedSalary))
    numerical_summary
    ```
